---
title: "Initial Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Personlized Bloodglucose Prediction

My journey in analyzing bloodglucose data of a type 1 diabetic patient, investigating the appropriate model design for this data and making predicitions based on this model/these models.

## Data

The data mainly concerns bloodglucose-levels, grams of carbohydrates consumed and insuline injections of one type 1 diabetic patient over the course of 6 months. For each event, the date, timestamp an a description of the even is available. Due to this, it is necessary to clean the data first. Below you can find the first few observations and a summary of the data before cleaning.

```{r, include=FALSE}
BGData = read.csv("Data BG.csv", header = TRUE)
```
```{r BGData}
head(BGData)
summary(BGData)
```
### Cleaning Data

We can see that the variable Event has many categories all measured in different ways. The category Carb (which indicates that the patient has eaten a meal that contained carbohydrates) is measured in grams (g) and can be easily seperated into its own variable. The latter is also true voor category Bloodglucose (bloodglucose value measured at a particular point of time in mmol/l).

The Carbs variable can be created by executing the following piece of code.

```{r}
Carbs = rep(NA, 3343)
Carbs[BGData$Event == "Carbs"] = BGData$Value[BGData$Event == "Carbs"]
head(data.frame(BGData$Event, BGData$Value, Carbs))
```

The last line allows us to check whether the transformation is performed correctly. Similarly, for Bloodglucose we execute

```{r}
Bloodglucose = rep(NA, 3343)
Bloodglucose[BGData$Event == "Bloodglucose"] = BGData$Value[BGData$Event == "Bloodglucose"]
head(data.frame(BGData$Event, BGData$Value, Bloodglucose))
```

Now we can transform the original variable Event and corresponding variable Value to discard the information on Carbs and Bloodglucose.

```{r}
BGData$Value[BGData$Event == "Carbs"] = NA
BGData$Value[BGData$Event == "Bloodglucose"] = NA
BGData$Event[BGData$Event == "Carbs"] = NA
BGData$Event[BGData$Event == "Bloodglucose"] = NA

head(BGData)
```

