---
title: "Initial Data Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Personlized Bloodglucose Prediction

My journey in analyzing bloodglucose data of a type 1 diabetic patient, investigating the appropriate model design for this data and making predicitions based on this model/these models.

## Data

The data mainly concerns bloodglucose-levels, grams of carbohydrates consumed and insuline injections of one type 1 diabetic patient over the course of 6 months. For each event, the date, timestamp an a description of the even is available. Due to this, it is necessary to clean the data first. Below you can find the first few observations and a summary of the data before cleaning.

```{r, include=FALSE}
BGData = read.csv("Data BG.csv", header = TRUE)
```
```{r BGData}
head(BGData)
summary(BGData)
```
### Cleaning Data

We can see that the variable Event has many categories all measured in different ways. The category Carb (which indicates that the patient has eaten a meal that contained carbohydrates) is measured in grams (g) and can be easily seperated into its own variable. The latter is also true voor category Bloodglucose (bloodglucose value measured at a particular point of time in mmol/l).

The Carbs variable can be created by executing the following piece of code.

```{r}
Carbs = rep(NA, 3343)
Carbs[BGData$Event == "Carbs"] = BGData$Value[BGData$Event == "Carbs"]
head(data.frame(BGData$Event, BGData$Value, Carbs))
```

The last line allows us to check whether the transformation is performed correctly. Similarly, for Bloodglucose we execute

```{r}
Bloodglucose = rep(NA, 3343)
Bloodglucose[BGData$Event == "Bloodglucose"] = BGData$Value[BGData$Event == "Bloodglucose"]
head(data.frame(BGData$Event, BGData$Value, Bloodglucose))
```

Now we can transform the original variable Event and corresponding variable Value to discard the information on Carbs and Bloodglucose.

```{r}
BGData$Value[BGData$Event == "Carbs"] = NA
BGData$Value[BGData$Event == "Bloodglucose"] = NA
BGData$Event[BGData$Event == "Carbs"] = NA
BGData$Event[BGData$Event == "Bloodglucose"] = NA
BGData$Event = factor(BGData$Event)
head(BGData)
```

Transforming the remainder of the categories in Event is a bit more complicated. The injection of insulin (bolus) is subdivided into multiple categories, which need to be extracted from the description. The various categories of this variable are

```{r}
levels(BGData$Event)
```

Three out of 10 categories are related to bolusses: Bolus, Combination bolus, Extended bolus. One seems to be related to insulin injection, but isn't called a bolus: Total amount insulin per day. The remainder of the categories pertain to other events related to the insulin pump (the device that regulates the insulin injection), pod (a tube placed into the pump that contains the insulin), some warnings and the change of time. Let's view the categories of the description variables.

```{r}
levels(BGData$Descripition1)
levels(BGData$Descripition2)
levels(BGData$Descripition3)
levels(BGData$Descripition4)
levels(BGData$Descripition5)
```

A quick analysis of the data together with the previous information leads to the following conclusions
1. Event "Total amount insulin per day" is a summary of the day's total insulin usage. In the description, this entry is broken up into two pieces of information: the "Bolus insulin" (total insulin injected directly as a result of eating or in order to correct for high bloodsugar) and the "Basal insulin" (insulin injection is spread over time).
2. Event "Bolus",  and "Combination bolus" mean that the pump was intructed to give insulin to the patient. While"Bolus" indicates injection pretty much immediately, with "Extended bolus" it will be injected slowly over a specified time window. A "Combination bolus" is the combination of these two options.

Due the aforementioned, the following variables are created

```{r}
ImmediateBolus = rep(NA, 3343)
ImmediateBolus[BGData$Descripition1 == "Direct dosis"&!is.na(BGData$Descripition1)] = BGData$ValueD1[BGData$Descripition1 == "Direct dosis"&!is.na(BGData$Descripition1)]
ExtendedBolus = rep(NA, 3343)
ExtendedBolus[BGData$Descripition2 == "Extended dosis"&!is.na(BGData$Descripition2)] = BGData$ValueD2[BGData$Descripition2 == "Extended dosis"&!is.na(BGData$Descripition2)]
DailyTotalBasalInsulin = rep(NA, 3343)
DailyTotalBasalInsulin[BGData$Descripition2 == "Basal insulin"&!is.na(BGData$Descripition2)] = BGData$ValueD2[BGData$Descripition2 == "Basal insulin"&!is.na(BGData$Descripition2)]
DailyTotalImmediateInsulin = rep(NA, 3343)
DailyTotalImmediateInsulin[BGData$Descripition1 == "Bolus insulin"&!is.na(BGData$Descripition1)] = BGData$ValueD1[BGData$Descripition1 == "Bolus insulin"&!is.na(BGData$Descripition1)]
BolusCancelled = rep(0, 3343)
BolusCancelled[BGData$Descripition2 == "Bolus cancelled"&!is.na(BGData$Descripition2)] = rep(1:3343)[BGData$Descripition2 == "Bolus cancelled"&!is.na(BGData$Descripition2)]
BolusCancelled[BGData$Descripition5 == "Bolus cancelled"&!is.na(BGData$Descripition5)] = rep(1:3343)[BGData$Descripition5 == "Bolus cancelled"&!is.na(BGData$Descripition5)]
BeforeMeal = rep(0, 3343)
BeforeMeal[BGData$Descripition1 == "Before meal"&!is.na(BGData$Descripition1)] = rep(1:3343)[BGData$Descripition1 == "Before meal"&!is.na(BGData$Descripition1)]
Occlusion = rep(0, 3343)
Occlusion[BGData$Descripition1 == "Occlusion found"&!is.na(BGData$Descripition1)] = rep(1:3343)[BGData$Descripition1 == "Occlusion found"&!is.na(BGData$Descripition1)]
ProposedBolus = rep(0, 3343)
ProposedBolus[BGData$Descripition1 == "Proposed bolus"&!is.na(BGData$Descripition1)] = BGData$ValueD1[BGData$Descripition1 == "Proposed bolus"&!is.na(BGData$Descripition1)]
ProposedBolus[BGData$Descripition2 == "Proposed bolus"&!is.na(BGData$Descripition2)] = BGData$ValueD2[BGData$Descripition2 == "Proposed bolus"&!is.na(BGData$Descripition2)]
ProposedBolus[BGData$Descripition4 == "Proposed bolus"&!is.na(BGData$Descripition4)] = BGData$ValueD4[BGData$Descripition4 == "Proposed bolus"&!is.na(BGData$Descripition4)]
DurationExtDos = rep(0, 3343)
DurationExtDos[BGData$Descripition1 == "Duration of Extended dosis"&!is.na(BGData$Descripition1)] = BGData$ValueD1[BGData$Descripition1 == "Duration of Extended dosis"&!is.na(BGData$Descripition1)]
DurationExtDos[BGData$Descripition3 == "Duration of Extended dosis"&!is.na(BGData$Descripition3)] = BGData$ValueD3[BGData$Descripition3 == "Duration of Extended dosis"&!is.na(BGData$Descripition3)]
```

This concludes the extraction of information for all bolus related events. The following dummy variables are created for the remaining events.

```{r}

PodAct = rep(0, 3343)
PodAct[BGData$Event == "Pod activated"] = 1
PodDeAct = rep(0, 3343)
PodDeAct[BGData$Event == "Pod deactivated"] = 1
PumpStart = rep(0, 3343)
PumpStart[BGData$Event == "Pump is started"] = 1
PumpStop = rep(0, 3343)
PumpStop[BGData$Event == "Time change (Winter time)"] = 1
TimeChange = rep(0, 3343)
TimeChange[BGData$Event == "Pod activated"] = 1
Warning  = rep(0, 3343)
Warning[BGData$Event == "Warning"] = 1

```

Combining all these newly created variables into a dateframe together with the day, date and timestamp results in the cleaned dataset.

```{r}
CleanedBGData = as.data.frame(cbind(BGData$DayOfWeek, BGData$Date, BGData$Time, BeforeMeal, Bloodglucose, BolusCancelled, Carbs, DailyTotalBasalInsulin, DailyTotalImmediateInsulin, DurationExtDos, ExtendedBolus, ImmediateBolus, Occlusion, PodAct, PodDeAct, ProposedBolus, PumpStart, PumpStop, TimeChange, Warning))
names(CleanedBGData)[1:3] = c("DayOfWeek", "Date", "Time")
save(CleanedBGData, file = "CleanedBGData.RData")
```
